---
title: "Phthalates panel, protein biomarkers (Empiria)"
author: "KC"
date: "4/26/2021"
output: html_document
---

Placental phthalates show AAPrev, urinary pthalates show AAPmono

lnAAPrev_nor: Summary placental phthalate measurement, normalized to placental weight (ln=ln-transformed) <br/>
lnAAPmonosg: Summary urinary phthalate measurement, normalized to specific gravity (ln=ln-transformed)

data: total-tides.csv, tides-fsex-3u-phthalates-plots.csv

# Placental

```{r echo=FALSE, fig.show="hold", message=FALSE, warning=FALSE, out.width="30%"}

library(readxl)
bo <- read_excel("/Users/kcarpio/Desktop/tides-fsex-phthalates-plots.xlsx")

bo$Tissue <- as.factor(bo$Tissue)
bo$BabySex <- as.factor(bo$BabySex)
bo$Sample <- as.factor(bo$Sample)

library(lme4)
library(lmerTest)
library(ggplot2)

# Placental phthalates

qplot(AAPrev_lnp, PPARg.empl, data=bo, color = Sample, linetype = BabySex,geom='point') +  
  geom_smooth(aes(group=interaction(BabySex)), method = "lm") + facet_grid(.~Tissue)+theme_minimal() + ggtitle("PPARg")+ theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + theme(plot.title = element_text(face = "bold"))

qplot(AAPrev_lnp, hCGa.empl, data=bo, color = Sample, linetype = BabySex,geom='point') +  
  geom_smooth(aes(group=interaction(BabySex)), method = "lm") + facet_grid(.~Tissue)+theme_minimal() + ggtitle("hCGa")+ theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + theme(plot.title = element_text(face = "bold"))

qplot(AAPrev_lnp, hCGb.empl, data=bo, color = Sample, linetype = BabySex,geom='point') +  
  geom_smooth(aes(group=interaction(BabySex)), method = "lm") + facet_grid(.~Tissue)+theme_minimal() + ggtitle("hCGb")+ theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + theme(plot.title = element_text(face = "bold"))

```

# Urinary, first trimester 

```{r echo=FALSE, fig.show="hold", message=FALSE, warning=FALSE, out.width="30%"}

library(readxl)
bo <- read_excel("/Users/kcarpio/Desktop/tides-fsex-phthalates-plots.xlsx")

bo$Tissue <- as.factor(bo$Tissue)
bo$BabySex <- as.factor(bo$BabySex)
bo$Sample <- as.factor(bo$Sample)

library(lme4)
library(lmerTest)
library(ggplot2)

# First trimester urinary

qplot(lnAAPmonosg, PPARg.empl, data=bo, color = Sample, linetype = BabySex,geom='point') +  
  geom_smooth(aes(group=interaction(BabySex)), method = "lm") + facet_grid(.~Tissue)+theme_minimal() + ggtitle("PPARg")+ theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + theme(plot.title = element_text(face = "bold"))

qplot(lnAAPmonosg, hCGa.empl, data=bo, color = Sample, linetype = BabySex,geom='point') +  
  geom_smooth(aes(group=interaction(BabySex)), method = "lm") + facet_grid(.~Tissue)+theme_minimal() + ggtitle("hCGa")+ theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + theme(plot.title = element_text(face = "bold"))

qplot(lnAAPmonosg, hCGb.empl, data=bo, color = Sample, linetype = BabySex,geom='point') +  
  geom_smooth(aes(group=interaction(BabySex)), method = "lm") + facet_grid(.~Tissue)+theme_minimal() + ggtitle("hCGb")+ theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + theme(plot.title = element_text(face = "bold"))

```

# Urinary, third trimester 

```{r echo=FALSE, fig.show="hold", message=FALSE, warning=FALSE, out.width="30%"}

library(readxl)
bo <- read_excel("/Users/kcarpio/Desktop/tides-fsex-3u-phthalates-plots.xlsx")

bo$Tissue <- as.factor(bo$Tissue)
bo$BabySex <- as.factor(bo$BabySex)
bo$Sample <- as.factor(bo$Sample)

library(lme4)
library(lmerTest)
library(ggplot2)

# Third trimester urinary

qplot(lnAAPmonosg_3U, PPARg.empl, data=bo, color = Sample, linetype = BabySex,geom='point') +  
  geom_smooth(aes(group=interaction(BabySex)), method = "lm") + facet_grid(.~Tissue)+theme_minimal() + ggtitle("PPARg")+ theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + theme(plot.title = element_text(face = "bold"))

qplot(lnAAPmonosg_3U, hCGa.empl, data=bo, color = Sample, linetype = BabySex,geom='point') +  
  geom_smooth(aes(group=interaction(BabySex)), method = "lm") + facet_grid(.~Tissue)+theme_minimal() + ggtitle("hCGa")+ theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + theme(plot.title = element_text(face = "bold"))

qplot(lnAAPmonosg_3U, hCGb.empl, data=bo, color = Sample, linetype = BabySex,geom='point') +  
  geom_smooth(aes(group=interaction(BabySex)), method = "lm") + facet_grid(.~Tissue)+theme_minimal() + ggtitle("hCGb")+ theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + theme(plot.title = element_text(face = "bold"))

```
